### 2019-11-04 paper/02mprelim.tex
Adding paper

\section{Mathematical preliminaries}
\subsection{Boundary value problems}
Consider a polygonal domain $\Omega \subset \mathbb{R}^2$ with boundary $\Gamma.$ 
Let $\gamma:[0,L] \to \Gamma$ be an arclength parametrization, and $\nu(t)$ denote the 
outward-pointing normal to $\Gamma$ at $\gamma(t).$  
Let $\omega \in \mathbb{C}$ in the first quadrant of the complex plane denote the 
Helmholtz wave number.
For ease of exposition our primary focus in the first portion of this paper will be on the 
following four boundary value problems.
\begin{enumerate}
\item The interior Dirichlet problem:
\begin{align}\label{eqn:BVPID}
(\Delta + \omega^2)u(x) &= 0, \hspace{1 cm} x \in \Omega,\\
u(\gamma(t)) &= f(t), \hspace{0.43 cm} t \in [0,L].
\end{align}

\item The exterior Dirichlet problem:
\begin{align}\label{eqn:BCPED}
(\Delta + \omega^2) u(x) &= 0, \hspace{1 cm} x \in \mathbb{R}^2\setminus\Omega,\\
u(\gamma(t)) &= f(t), \hspace{0.43 cm} t \in[0,L].
\end{align}

\item The interior Neumann problem:
\begin{align}\label{eqn:BVPIN}
(\Delta + \omega^2) u(x) &= 0, \hspace{1 cm} x \in \Omega,\\
\nabla u(\gamma(t))\cdot \nu(t)  &= f(t), \hspace{0.45 cm} t\in [0,L].
\end{align}

\item The exterior Neumann problem:
\begin{align}\label{eqn:BVPEN}
(\Delta + \omega^2) u(x) &= 0, \hspace{1 cm} x \in \mathbb{R}^2\setminus\Omega,\\
\nabla u(\gamma(t))\cdot \nu(t)  &= f(t), \hspace{0.45 cm} t\in [0,L].
\end{align}

\end{enumerate}

\begin{remark}
The existence and uniqueness of solutions to the above equations is a classical result. 
The exterior Dirichlet problem and interior Neumann problem require the additional assumption that
$$\int_0^L f(t)\,{\rm d}t = 0.$$
\end{remark}