### 2019-11-05 paper/02mprelim.tex

\section{Mathematical preliminaries}
\subsection{Boundary value problems}
Consider a polygonal domain $\Omega \subset \mathbb{R}^2$ with boundary $\Gamma.$ 
Let $\gamma:[0,L] \to \Gamma$ be an arclength parametrization, and $\nu(t)$ denote the 
outward-pointing normal to $\Gamma$ at $\gamma(t).$  
Let $\omega \in \mathbb{C} \setminus \{ 0 \}$ in the first quadrant of the complex plane denote the 
Helmholtz wave number.
\begin{enumerate}
\item The interior Dirichlet problem for Laplace's equation:
\begin{align}\label{eqn:BVPIDL}
\Delta u(\bx) &= 0, \hspace{1 cm} \bx \in \Omega,\\
u(\gamma(t)) &= f(t), \hspace{0.43 cm} t \in [0,L].
\end{align}

\item The interior Dirichlet problem for Helmholtz's equation:
\begin{align}\label{eqn:BVPIDH}
(\Delta + \omega^2) u(\bx) &= 0, \hspace{1 cm} \bx \in \Omega,\\
u(\gamma(t)) &= f(t), \hspace{0.43 cm} t \in [0,L].
\end{align}

\item The exterior Dirichlet problem for Laplace's equation:
\begin{align}\label{eqn:BVPEDL}
\Delta u(\bx) &= 0, \hspace{1 cm} \bx \in \mathbb{R}^2\setminus\Omega,\\
u(\gamma(t)) &= f(t), \hspace{0.43 cm} t \in[0,L], \\
u(\bx) &= O(1) , \hspace{0.4cm} |\bx| \to \infty.
\end{align}

\item The exterior Dirichlet problem for Helmholtz's equation:
\begin{align}\label{eqn:BVPEDH}
\Delta u(\bx) &= 0, \hspace{1 cm} x \in \mathbb{R}^2\setminus\Omega,\\
u(\gamma(t)) &= f(t), \hspace{0.43 cm} t \in[0,L], \\
\sqrt{|\bx|} \left|\partial_{|\bx|} u - i\omega u(\bx)\right| &\to 0 , \hspace{0.86cm} |\bx| \to \infty.
\end{align}

\item The interior Neumann problem for Laplace's equation:
\begin{align}\label{eqn:BVPINL}
\Delta u(\bx) &= 0, \hspace{1 cm} \bx \in \Omega,\\
\nabla u(\gamma(t))\cdot \bnu(t)  &= f(t), \hspace{0.45 cm} t\in [0,L],\\
\int_{0}^{L} f(t) dt &= 0.
\end{align}

\item The interior Neumann problem for Helmholtz's equation:
\begin{align}\label{eqn:BVPINH}
(\Delta + \omega^2) u(\bx) &= 0, \hspace{1 cm} \bx \in \Omega,\\
\nabla u(\gamma(t))\cdot \bnu(t)  &= f(t), \hspace{0.45 cm} t\in [0,L].
\end{align}

\item The exterior Neumann problem for Laplace's equation:
\begin{align}\label{eqn:BVPEDL}
\Delta u(\bx) &= 0, \hspace{1 cm} \bx \in \mathbb{R}^2\setminus\Omega,\\
\nabla u(\gamma(t))\cdot \bnu(t)  &= f(t), \hspace{0.43 cm} t \in[0,L], \\
\left|u(\bx) + \left(\frac{1}{2\pi} \int_{0}^{L} f(t) dt\right) \log{|\bx|}\right| &\to 0 , \hspace{0.4cm} |\bx| \to \infty.
\end{align}

\item The exterior Neumann problem for Helmholtz's equation:
\begin{align}\label{eqn:BVPEDH}
\Delta u(\bx) &= 0, \hspace{1 cm} x \in \mathbb{R}^2\setminus\Omega,\\
\nabla u(\gamma(t))\cdot \bnu(t)  &= f(t), \hspace{0.43 cm} t \in[0,L], \\
\sqrt{|\bx|} \left|\partial_{|\bx|} u - i\omega u(\bx)\right| &\to 0 , \hspace{0.86cm} |\bx| \to \infty.
\end{align}

\end{enumerate}

\begin{remark}
The existence and uniqueness of solutions to the above equations is a classical result. 
The interior Dirichlet and Neumann problems for Helmholtz equation have a countable set of
real $\omega$'s (which only accumulate at $\infty$)  for which the PDE has a non-unique solution. 
\end{remark}